---
title: "Final"
author: "Kai Itokazu"
subtitle: MGSC 310 Midterm
output:
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

library(knitr)

# Change the number in set seed to your own favorite number
set.seed(1818)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = FALSE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```

```{r setup-2}

# Always print this out before your assignment
sessionInfo()
getwd()

```

```{r setup-3}

# load all your libraries in this chunk 
library('tidyverse')
# note, do not run install.packages() inside a code chunk. install them in the console outside of a code chunk. 

```


```{r}

library('here')
nba_players <-
  read.csv(here("FINAL","salary_player.csv"))

```

```{r}
library("nbastatR")
library('ggplot2')
library('highcharter')
library('dplyr')
library(dplyr)
library(magrittr)
library(priceR)

library(shiny)
library(shinythemes)



```

### Data Manipulation



#### Filtering
```{r, results='hide'}
nba_players<-
  nba_players %>% mutate(raptor_score = Raptor...)

#Removing statistically insignificant players and adjusting for inflation
nba_players %<>% filter(MPG > 10, G > 20) %>% 
  mutate(salary_adjusted = adjust_for_inflation(salary,season_start,"US", to_date = max(year_id)),
         salary_M = salary_adjusted / 1000000,
         WAR_per_million = Raptor.WAR/salary_M,
         Raptor_per_million = raptor_score/salary_M)
```

### MORE DATA CLEANING
```{r}
library(here)
library(dplyr)
library(magrittr)
library(priceR)
library(highcharter)
library(hash)
library(stringr)
library(forcats)

players_historic <- read.csv(here("FINAL/FINAL_DATASETS/NBA_Players.csv"))
salaries <- read.csv(here("FINAL/FINAL_DATASETS/salaries_1985to2018.csv"))

glimpse(players_historic)

players_historic %<>% rename(Raptor_T = Raptor...,
                             Raptor_O = Raptor.O,
                             Raptor_D = Raptor.D,
                             Raptor_WAR = Raptor.WAR) %>% 
  filter(type == "RS", Min > 50) %>% 
  arrange(age) %>% arrange(player_id) %>% 
  select(player_id, name_common, age, pos, year_id, team_id, G, Min, MPG, starts_with("Raptor")) %>% 
  mutate(team_id = fct_collapse(team_id, CHA = c("CHA", "CHH"), WAS = c("WAS", "WSB")))

salaries %<>% select(player_id, salary, season_end)

players <- players_historic %>% 
  inner_join(salaries, by = "player_id", "year_id") %>% 
  filter(year_id == season_end,) %>%
  select(!season_end)

players %<>% group_by(player_id, year_id) %>% 
  filter(n() == 1) %>% 
  arrange(year_id) %>% 
  arrange(player_id)

players$salary_adjusted <- c(adjust_for_inflation(players$salary, players$year_id, "US", to_date = max(players$year_id)))
players %<>% mutate(salary_M = round(salary_adjusted / 1000000, 3),
                   Raptor_per_million = Raptor_T/salary_M,
                   WAR_per_million = Raptor_WAR/salary_M)

players_clean <- players %>% group_by(player_id) %>% 
  mutate(Raptor_F = lead(Raptor_T, order_by = year_id),
         Raptor_WAR_F = lead(Raptor_WAR, order_by = year_id),
         salary_M_F = lead(salary_M, order_by = year_id),
         WAR_per_million_F = lead(WAR_per_million, order_by = year_id),
         Raptor_per_million_F = lead(Raptor_per_million, order_by = year_id),
         age_squared = (age)^2) %>%
  filter(!is.na(Raptor_F))
players_clean



```





### Data Summary

#### Summary Statistics by season 1985-2018
```{r}
season_summary <- nba_players %>% group_by(year_id) %>% 
  summarise(salary_sum = sum(salary_M),
            salary_ave = mean(salary_M),
            player_count = str_count(toString(player_id)),
            raptor_ave = mean(raptor_score))
season_summary
```
#### Average and sum of slaries per season 1985-2018
```{r}
ggplot(data = season_summary, aes(x = year_id, y = salary_sum)) + geom_col() #1987 doesn't follow pattern
ggplot(data = season_summary, aes(x = year_id, y = salary_ave)) + geom_col() #1990 doesn't follow pattern
```


```{r}
season_summary %>% filter(player_count < 2000) #Verifying which years have insufficient data
nba_players_clean <- nba_players %>% filter(year_id == 1987, 
                                            year_id == 1990)#Filtering out seasons without sufficient data for seasonal data analysis
```


### Plots

#### Plot of average RAPTOR rating and average salary (adjusted for inflation) through each NBA players career
```{r, message=FALSE}
# ggplot(data = nba_players, aes(x=raptor_score, y = salary_adjusted)) + geom_point() + xlim


average_player_stats <- nba_players %>% 
  group_by(player_id, name_common) %>%
  summarise(across(where(is.numeric), mean, na.rm = FALSE))

#Minutes per Game vs Raptor Score
ggplot(data = average_player_stats, aes(x=raptor_score, y = salary_adjusted)) + geom_point(alpha = 0.5)  + ggtitle("Average Player Stats - Avg Raptor Score vs Average Salary (adjusted for inflation) - 1985-2018") + xlim(-10, 10) + labs(x = "Average Raptor Score", y = "Average Salary")



```

#### Average Raptor WAR vs Average Salary (adjusted inflation) over the years. Players with Raptor Scores over 0
```{r}
average_player_stats <- nba_players %>% 
  group_by(player_id, name_common) %>%
  summarise(across(where(is.numeric), mean, na.rm = FALSE))

#Filter players that are greater than 0
players_score_positive <- average_player_stats %>% 
  filter(raptor_score >= 0)

#ROUND
players_score_positive <- players_score_positive %>% mutate(raptor_war = round(Raptor.WAR,2))
players_score_positive <- players_score_positive %>% mutate(salary_M = round(salary_M,2))

hchart(players_score_positive, "scatter", hcaes(x="Raptor.WAR", y="salary_M", group="name_common")) %>% 
  hc_tooltip(pointFormat = "<b>{point.name_common}</b><br/>AVG RAPTOR WAR Score:{point.raptor_war}<br/><br/>AVG Salary: {point.salary_M}<br/>") %>%
  hc_title(text="Raptor WAR vs Salary") %>% 
  hc_subtitle(text="Average WAR Score vs Average Salary in Millions (adjusted inflation) over the years. Players with Raptor Scores over 0") %>% 
  hc_legend(enabled = FALSE)




```


#### Raptor score and salary based on 2018
```{r}

library('highcharter')
library('dplyr')
nba_players_2018 <- nba_players %>% 
  filter(season_end == 2018)

nba_players_2018<-
  nba_players_2018 %>% mutate(raptor_score = Raptor...)


hchart(nba_players_2018, "scatter", hcaes(x="raptor_score", y="salary", group="team")) %>% 
  hc_tooltip(pointFormat = "<b>{point.name_common}</b><br/>Raptor Score:{point.raptor_score}<br/><br/>Salary: {point.salary}<br/>") %>%
  hc_title(text="Raptor Score vs Salary") %>% 
  hc_subtitle(text="2018 Season Data")

```



#### Raptor score and salary based on 2018 - GROUPED BY POSITIONS
```{r}
nba_players_2018 <- nba_players %>% 
  filter(season_end == 2018)

nba_players_2018<-
  nba_players_2018 %>% mutate(raptor_score = Raptor...)

nba_players_2018<-
  nba_players_2018 %>% mutate(salary_M = round(salary_M,2))


hchart(nba_players_2018, "scatter", hcaes(x="raptor_score", y="salary_M", group="pos")) %>% 
  hc_tooltip(pointFormat = "<b>{point.name_common}</b><br/>Raptor Score:{point.raptor_score}<br/><br/>Salary: {point.salary_M}<br/>") %>%
  hc_title(text="Raptor Score vs Salary (By Position)") %>% 
  hc_subtitle(text="2018 Season Data")




```


#### Plot of average RAPTOR rating and average minutes per game through each NBA players career
```{r, message=FALSE}
average_player_stats <- nba_players %>% 
  group_by(player_id, name_common) %>%
  summarise(across(where(is.numeric), mean, na.rm = FALSE))

#Minutes per Game vs Raptor Score
ggplot(data = average_player_stats, aes(x=raptor_score, y = MPG)) + geom_point(alpha = 0.5)  + ggtitle("Average Player Stats - Avg Raptor Score vs Average Minutes Per Game  (1985-2018) ") + xlim(-10, 10) + labs(x = "Average Raptor Score", y = "Average Minutes Per Game")
```


### PREDICTING A PLAYER'S SALARY FOR THE FOLLWING SEASON



```{r}
library(shiny)
library(shinythemes)
library(data.table)
library(RCurl)
library(randomForest)

rf_mods <- list()
mse_err <- NULL
test_err <- NULL

#THIS TAKES FOREVER
#Hyperparameter Tuning
# for(mtry in 1:9){
#   rt_fit <- randomForest(salary_M_F ~ MPG + Raptor_T + Raptor_WAR + age + factor_pos + salary_M, 
#                          data = players_clean, 
#                          type = regression, 
#                          ntree = 500, 
#                          mtry = mtry, 
#                          importance = TRUE)
#   
#   mse_err[mtry] <- rt_fit$mse[500]
# 
#   cat(mtry," ")
# }
# 
# mse_err
# 
# res_DF <- data.frame(mtry = 1:9, mse_err)
# 
# ggplot(res_DF, aes(x = mtry, y = mse_err)) + geom_point() + theme_minimal()


# players_clean <- players_clean %>% mutate(factor_pos = as.factor(pos))
glimpse(players_clean) 

players_clean <- players_clean %>% mutate(factor_pos = as.factor(pos))
#Random Forest
rt_fit <- randomForest(salary_M_F ~ MPG + Raptor_T + Raptor_WAR + age + factor_pos + salary_M, data = players_clean, type = regression, ntree = 500, mtry = 3, importance = TRUE)


#Linear Regression Model
lm_mod <- lm(salary_M_F ~ MPG + Raptor_T + Raptor_WAR + age + factor_pos + salary_M, data = players_clean)


#HISTOGRAM FOR HELP
ggplot(data = nba_players, aes(x = salary_M)) + 
  geom_histogram() + ggtitle("Salary Distribution") +
  xlab("Salary (in Millions)") + ylab("Frequency")


glimpse(nba_players)

max(nba_players$age)
min(nba_players$age)

factor(nba_players$pos)

print(rt_fit) 

plot(rt_fit)

#variable importance plot
varImpPlot(rt_fit, type = 1, scale = FALSE)

#Importance of Plot
importance(rt_fit, type=1, scale = FALSE)

#Mininum Depth Distribution
# plot_min_depth_distribution(rt_fit)

# explain_forest(rt_fit, 
#                interactions = TRUE, 
#                data = players_clean)


#
# --------- R SHINY DASHBOARD -----------
#
#---- USER INTERFACE --------
ui <- fluidPage(theme = shinytheme("united"),
  
  # Page header
  headerPanel('Estimating Salary (in Millions)'),
  
  # Input values
  sidebarPanel(
    
    HTML("<h3>Select Position</h3>"),
    
    selectInput("factor_pos", label = "Position:", 
                choices = list("Center" = "C",
                               "Power Forward" = "PF", 
                               "Point Guard" = "PG", 
                               "Small Forward" = "SF",
                               "Shooting Guard" = "SG"), 
                selected = "Power Forward"),
    
    
    sliderInput("salary_M", label = "Current Season Salary:",
                min = 0, max = 50,
                value = 2),
    
    sliderInput("MPG", label = "Minutes Per Game:",
                min = 10, max = 40,
                value = 24),
    
    sliderInput("Raptor_T", label = "Raptor Score:",
                min = -14, max = 14,
                value = 0),
    
    sliderInput("Raptor_WAR", label = "Raptor Wins Above Replacement:",
                min = -5, max = 24,
                value = 0),

    sliderInput("age", label = "Age",
                min = 18, max = 42,
                value = 25),
    
    actionButton("submitbutton", "Submit", class = "btn btn-primary")
  ),
  
  mainPanel(
    tags$label(h3('Output')), # Status/Output Text Box
    verbatimTextOutput('contents'),
    tableOutput('tabledata') # Prediction results table
    
  )
)



#------- SERVER -------
server <- function(input, output, session) {

  # Input Data
  datasetInput <- reactive({

  df <- data.frame(
    Name = c("factor_pos","salary_M","MPG","Raptor_T","Raptor_WAR","age"),
    Value = as.character(c(input$factor_pos,
                           input$salary_M,
                           input$MPG,
                           input$Raptor_T,
                           input$Raptor_WAR,
                           input$age)),
    stringsAsFactors = FALSE)

  sal <- "salary_M"
  df <- rbind(df, sal)
  input <- transpose(df)
  write.table(input,"input.csv", sep=",", quote = FALSE, row.names = FALSE, col.names = FALSE)
  
  test <- read.csv(paste("input", ".csv", sep=""), header = TRUE)
  
  test$factor_pos <- factor(test$factor_pos, levels = c("C","PF","PG","SF","SG"))
  
  Output <- data.frame(RandomForest=round(predict(rt_fit,test,type="response"), 2),
                       LinearRegression=round(predict(lm_mod,test,type="response"), 2))
  print(Output)
  
  
  
})
  
  # Status/Output Text Box
  output$contents <- renderPrint({
    if (input$submitbutton>0) { 
      isolate("Estimated Player Salary") 
    } else {
      return("Enter Player Information")
    }
  })
  
  # Prediction results table
  output$tabledata <- renderTable({
    if (input$submitbutton>0) { 
      isolate(datasetInput()) 
    } 
  })
  
}

####################################
# Create the shiny app             #
####################################
shinyApp(ui = ui, server = server) 

```

